---
title: "FDI project"
format: html
---

```{r}
library(tidyverse)

set.seed(101)

advanced_macro_data <- expand_grid(
  country = c("China", "South Korea", "United States", "Japan", "Taiwan"),
  year = 1980:2025
) %>%
  mutate(
    # Simulate Foreign Direct Investment with some variance
    fdi_bln = case_when(
      country == "China" & year > 2000 ~ (year - 2000)^1.5 * 10 + rnorm(n(), 50, 20),
      country == "United States" ~ (year - 1980) * 12 + rnorm(n(), 100, 30),
      TRUE ~ (year - 1980) * 5 + rnorm(n(), 20, 10)
    ),
    # Simulate Unemployment Rate (spiking during known crises)
    unemployment_rate = case_when(
      country == "South Korea" & year %in% 1997:1999 ~ rnorm(n(), 7.5, 1),
      country == "United States" & year %in% c(2008, 2009, 2010, 2020) ~ rnorm(n(), 8.5, 1.5),
      TRUE ~ rnorm(n(), 4.0, 1.2)
    ),
    # Simulate Inflation - introducing NA values randomly to force data cleaning
    inflation_rate = rnorm(n(), 3.0, 2.5),
    # Simulate Government Debt as % of GDP
    gov_debt_pct = case_when(
      country == "Japan" ~ (year - 1980) * 4 + 50 + rnorm(n(), 10, 5),
      country == "United States" & year > 2000 ~ (year - 2000) * 3 + 60 + rnorm(n(), 5, 5),
      TRUE ~ (year - 1980) * 1.5 + 20 + rnorm(n(), 5, 5)
    ),
    # Simulate Tech Exports
    tech_exports_bln = if_else(
      country %in% c("South Korea", "Taiwan", "China") & year > 1995,
      (year - 1995)^2 * 1.2 + rnorm(n(), 10, 5),
      (year - 1980) * 2 + rnorm(n(), 5, 2)
    ),
    # Add a messy text column mimicking qualitative analyst notes
    analyst_note = sample(
      c(" stable_growth ", "CRISIS_watch", "recession Warning", " steady ", NA), 
      size = n(), replace = TRUE, prob = c(0.4, 0.1, 0.1, 0.3, 0.1)
    )
  ) %>%
  # Introduce random missing values (NAs) in inflation to practice cleaning
  mutate(inflation_rate = ifelse(runif(n()) < 0.1, NA, inflation_rate)) %>%
  # Round numerics
  mutate(across(where(is.numeric) & !matches("year"), ~round(.x, 2)))

# Save to your working directory
write_csv(advanced_macro_data, "advanced_macro_data.csv")
```

# Question 1
# Calculate the all-time average and standard deviation of fdi_bln for each country. Your final output should be a small table with three columns: the country, the average FDI, and the standard deviation. Sort this table so the country with the highest average FDI is at the top.

```{r}
advanced_macro_data %>%
    group_by(country) %>%
    summarize(
        avg_fdi = mean(fdi_bln, na.rm = TRUE),
        sd_fdi = sd(fdi_bln, na.rm = TRUE)
    ) %>%
    arrange(desc(avg_fdi))
```

# Question 1.2
# Calculate the average and median gov_debt_pct for each country in advanced_macro_data. Output a table with three columns: the country, average debt, and median debt. Sort the output so the nation with the lowest median debt appears at the top.

```{r}
advanced_macro_data %>%
    group_by(country) %>%
    summarize(
        avg_debt = mean(gov_debt_pct, na.rm = TRUE),
        med_debt = median(gov_debt_pct, na.rm = TRUE)
    ) %>%
    arrange(desc(med_debt))
```

# Question 2
# Find the maximum unemployment_rate recorded for each country, and the specific year that maximum occurred. Hint: You want to filter rows based on a group maximum, not just summarize.

```{r}
advanced_macro_data %>%
    group_by(country) %>%
    filter(unemployment_rate == max(unemployment_rate, na.rm = TRUE)) %>%
    select(country, year, unemployment_rate)
```

# Question 2.2
# Find the specific year each country recorded its peak fdi_bln value. Output a table containing only the country, year, and fdi_bln columns, sorted so the overall highest FDI peak across all nations appears at the top.

```{r}
advanced_macro_data %>%
    group_by(country) %>%
    filter(fdi_bln == max(fdi_bln, na.rm = TRUE)) %>%
    select(country, year, fdi_bln) %>%
    arrange(desc(fdi_bln))
```

# Question 3 
# Group the data by country. In a single pipeline, count how many missing values (NA) exist in the inflation_rate column for each nation.

```{r}
advanced_macro_data %>%
    group_by(country) %>%
    summarize(missing_inflation_count = sum(is.na(inflation_rate))) %>%
    arrange(desc(missing_inflation_count))
```

# Question 3.2
# Group the data by country. In a single pipeline, count how many missing values (NA) exist in the analyst_note column for each nation, and sort the output so the country with the most missing notes is at the top.

```{r}
advanced_macro_data %>%
    group_by(country) %>%
    summarize(missing_analyst_note = sum(is.na(analyst_note))) %>%
    arrange(desc(missing_analyst_note))
```

# Question 4
# Create a new column called decade (e.g., 1980, 1990, 2000). Then, calculate the median gov_debt_pct for each country, per decade.
```{r}
advanced_macro_data %>%
    mutate(decade = year %/% 10 * 10) %>%
    group_by(country, decade) %>%
    summarize(
        median_debt = median(gov_debt_pct, na.rm = TRUE)
    )
```

# Question 4.2
# Create a new column called five_year_block (e.g., 1980, 1985, 1990). Then, calculate the average unemployment_rate for each country in each 5-year block. Hint: Use modulo or integer division with 5 instead of 10.

```{r}
advanced_macro_data %>%
    mutate(five_year_block = year %/% 5 * 5) %>%
    group_by(country, five_year_block) %>%
    summarize(
        avg_unemployment = mean(unemployment_rate, na.rm = TRUE)
    )
```

# Question 5
# For South Korea only, create a new column that shows the inflation_rate of the previous year. Then, create another column calculating the year-over-year change in inflation (Current Year - Previous Year)

```{r}
advanced_macro_data %>%
    filter(country == 'South Korea') %>%
    arrange(year) %>%
    mutate(
        prev_inflation = lag(inflation_rate),
        yoy_inflation_change = inflation_rate - prev_inflation
    )
```

# Question 5.2
# Filter the dataset for the "United States" only. Create a new column showing the gov_debt_pct of the previous year, and another column showing the year-over-year change in debt percentage (Current Year - Previous Year).

```{r}
advanced_macro_data %>%
    filter(country == 'United States') %>%
    arrange(year) %>%
    mutate(ly_gov_debt_pct = lag(gov_debt_pct),
    yoy_gov_debt_pct = gov_debt_pct - ly_gov_debt_pct
    ) %>%
    select(country, year, gov_debt_pct, ly_gov_debt_pct, yoy_gov_debt_pct)
```
# Learned how to use lag functions and properly ensure newly mutated columns appear in the pulled dataset.

# Question 6
# Calculate the cumulative sum of tech_exports_bln for Taiwan from 1980 to 2025. Plot this trajectory as an area chart (geom_area), showing how their total tech export volume has accumulated over time.

```{r}
advanced_macro_data %>%
    filter(country == 'Taiwan') %>%
    arrange(year) %>%
    mutate(total_exports = cumsum(tech_exports_bln)) %>%
    ggplot(aes(x = year, y = total_exports)) +
    geom_area()
```

# learned the cumsum aggregate function and how summarize will collapse the dataset into one group which is the same no matter the other details whereas mutate will not collapse it allowing it to make a cumulative sum that summarize cannot do.

# Question 6.2
# Filter the dataset for South Korea and Japan for years between 2000 and 2025. Calculate the cumulative sum of Foreign Direct Investment (fdi_bln) for each country over this period. Plot these two trajectories on the same chart as time-series lines (geom_line()), coloring the lines by country to compare their total accumulated FDI.

```{r}
advanced_macro_data %>%
filter(country %in% c('Japan', 'South Korea'), year >= 2000) %>%
group_by(country) %>%
arrange(year) %>%
mutate(sum_fdi_bln = cumsum(fdi_bln)) %>%
ggplot(aes(x = year, y = sum_fdi_bln, color = country)) +
geom_line()
```

# Question 6.3
# Filter the dataset for China, Taiwan, and South Korea for years from 1990 to 2025. Calculate the running cumulative total of technology exports (tech_exports_bln) for each country over this time frame. Plot their growth on a single line chart (geom_line()), coloring each line by country to compare how their high-tech export bases accumulated over those 35 years.

```{r}
advanced_macro_data %>%
    filter(country %in% c('China', 'Taiwan', 'South Korea'), year >= 1990) %>%
    arrange(year) %>%
    mutate(sum_exports = cumsum(tech_exports_bln)) %>%
    ggplot(aes(x = year, y = sum_exports, color = country)) +
    geom_line()
```

# Learned about %in% and properly using it which means anything in the data set that fulfills any of the rules within, and then how to use c() which means combine, by using arrange(year) we ensure the values do not mix together from all of the entries inside the c() work.

# Question 6.4
# Filter the dataset for United States, Japan, and China for years from 2005 to 2025. Calculate the cumulative sum of gov_debt_pct for each country across this timeframe. Plot the trajectories on a line chart (geom_line()), ensuring each country has its own distinct line color.

```{r}
advanced_macro_data %>%
    filter(country %in% c('United States', 'China', 'Japan'), year >= 2005) %>%
    group_by(country) %>%
    arrange(year) %>%
    mutate(sum_gdebtp = cumsum(gov_debt_pct)) %>%
    ggplot(aes(x = year, y = sum_gdebtp, color = country)) +
    geom_line()
```
# Learned I must group by countries otherwise it will combine their rates in the final graph


# Question 7
# Let's smooth out the volatility. Calculate a 3-year rolling average for China's fdi_bln. Plot both the actual yearly FDI (as points) and the 3-year rolling average (as a line) on the same graph. Hint: Look up the slider or zoo packages, or figure out how to do it manually with dplyr lags.

```{r}
advanced_macro_data %>%
    filter(country == 'China') %>%
    arrange(year) %>%
    mutate(
        fdi_rolling_avg = ((fdi_bln + lag(fdi_bln) + lag(fdi_bln, 2)) /3)
    ) %>%
    ggplot(aes(x = year)) +
    geom_point(aes(y = fdi_bln), color = 'black') +
    geom_line(aes(y = fdi_rolling_avg), color = 'blue', linewidth = 1)
```

# Question 7.1
# Filter the dataset for South Korea. Calculate a 3-year rolling average for its technology exports (tech_exports_bln). Create a visualization that plots the actual annual tech exports as points (geom_point()) and the smoothed 3-year rolling average as a line (geom_line()) on top of it.

```{r}
advanced_macro_data %>%
    filter(country == 'South Korea') %>%
    arrange(year) %>%
    mutate(
        rol_avg_tech = ((tech_exports_bln + lag(tech_exports_bln) + lag(tech_exports_bln, 2)) / 3)
    ) %>%
    ggplot(aes(x = year)) +
    geom_point(aes(y = tech_exports_bln), color = 'purple') +
    geom_line(aes(y = rol_avg_tech), color = 'orange')
```

# Question 7.2
# Filter the dataset for the United States for years between 1990 and 2025. Calculate a 4-year rolling average for its government debt ratio (gov_debt_pct). Plot the raw yearly debt percentages as points and the 4-year rolling average trend as a line on the same graph.

```{r}
advanced_macro_data %>%
    filter(country == 'United States', year >= 2000) %>%
    arrange(year) %>%
    mutate(
        rol_avg_gdpt = ((gov_debt_pct) + lag(gov_debt_pct) + lag(gov_debt_pct, 2) + lag(gov_debt_pct, 3)) / 4
    ) %>%
    ggplot(aes(x = year)) +
    geom_point(aes(y = gov_debt_pct, color = 'blue')) +
    geom_line(aes(y = rol_avg_gdpt, color = 'red'))
```

# Question 8
# Pivot the dataset so that each country's tech_exports_bln has its own dedicated column (e.g., your columns should be year, China, South Korea, United States, etc.). This is known as making data "wide".

```{r}
advanced_macro_data %>%
    select(year, country, tech_exports_bln) %>%
    pivot_wider(
        names_from = country,
        values_from = tech_exports_bln
    )
```

# Question 8.1
# Reshape the dataset to analyze Foreign Direct Investment (fdi_bln) across nations side by side.
# Select only year, country, and fdi_bln. Pivot the data so that each country becomes its own column filled with its respective FDI numbers.
```{r}
advanced_macro_data %>%
    select(year, country, fdi_bln) %>%
    pivot_wider(
        names_from = country,
        values_from = fdi_bln
    )
```
# Question 8.2
# Create a wide table comparing Government Debt (gov_debt_pct) for only the Asian economies in the dataset (China, Japan, South Korea, and Taiwan) for years starting from 2000.
# Filter for those four countries and year >= 2000. Select year, country, and gov_debt_pct. Pivot the dataset wide so each of those four countries gets its own column.

```{r}
advanced_macro_data %>%
    filter(country %in% c('China', 'Japan', 'Taiwan', 'South Korea')) %>%
    select(year, country, gov_debt_pct) %>%
    pivot_wider(
        names_from = country,
        values_from = gov_debt_pct
    )
```

# Question 8.3
# Calculate a 3-year rolling average for Unemployment Rate (unemployment_rate) for United States and China, then pivot the calculated rolling averages wide.
# Filter for United States and China. Group by country and order chronologically with arrange(year). Create a new column unemp_rolling_avg using lag(). Select year, country, and unemp_rolling_avg.
# Pivot wide so United States and China are column headers displaying their smoothed unemployment rates.

```{r}
advanced_macro_data %>%
    filter(country %in% c('United States', 'China')) %>%
    arrange(year) %>%
    mutate(rol_avg = ((unemployment_rate) + lag(unemployment_rate) + lag(unemployment_rate, 2)) / 3) %>%
    select(year, country, rol_avg) %>%
    pivot_wider(
        names_from = country,
        values_from = rol_avg
    )
```

# Question 9
# Using the wide data you created in Question 8, calculate a new column representing the difference between South Korea's tech exports and Taiwan's tech exports for every year. Plot this difference as a bar chart over time.
# Q8
```{r}
advanced_macro_data %>%
    select(year, country, tech_exports_bln) %>%
    pivot_wider(
        names_from = country,
        values_from = tech_exports_bln
    ) %>%
    mutate(tech_diff = `South Korea` - `Taiwan`) %>%
    ggplot(aes(x = year, y = tech_diff)) +
    geom_col(fill = 'blue') +
    labs(
        title = "Tech Export Difference: South Korea and Taiwan",
        x = "Year",
        y = "Difference (Billions USD)"
    )
```

# Question 9.1
# Compare the Unemployment Rate (unemployment_rate) gap between the United States and Japan. Pivot the dataset wide, create a new column unemp_gap (US rate minus Japan's rate), and plot this gap over time as a bar chart (geom_col()).
```{r}
advanced_macro_data %>%
    select(year, country, unemployment_rate) %>%
    pivot_wider(
        names_from = country,
        values_from = unemployment_rate
    ) %>%
    mutate(unemp_gap = `United States` - `Japan`) %>%
    ggplot(aes(x = year, y = unemp_gap)) +
    geom_col(fill = 'red') +
    labs(
        title = "Unemployment Gap between the United States and Japan",
        x = "Year",
        y = "Unemployment Gap (In Percent)"
    )
```

# Question 9.2
# Calculate the difference in Foreign Direct Investment (fdi_bln) between China and South Korea for years starting from 2005. Pivot the data wide, create a new column fdi_diff (China minus South Korea), and plot the FDI difference over time as a bar chart (geom_col()).

```{r}
advanced_macro_data %>%
    filter(year >= 2005) %>%
    select(year, country, fdi_bln) %>%
    pivot_wider(
        names_from = country,
        values_from = fdi_bln
    ) %>%
    mutate(fdi_diff = `China` - `South Korea`) %>%
    ggplot(aes(x = year, y = fdi_diff)) +
    geom_col(fill = 'purple') +
    labs(
        title = "FDI difference between China and South Korea",
        x = "Year",
        y = "FDI difference"
    )
```



# Question 9.3
# Calculate the combined sum of Government Debt (gov_debt_pct) ratios for Taiwan and South Korea. Pivot the data wide, create a new column total_debt_pct (Taiwan plus South Korea), and plot this combined total over time as a bar chart (geom_col()).

```{r}
advanced_macro_data %>%
    select(year, country, gov_debt_pct) %>%
    pivot_wider(
        names_from = country,
        values_from = gov_debt_pct
    ) %>%
    mutate(total_debt_pct = `Taiwan` + `South Korea`) %>%
    ggplot(aes(x = year, y = total_debt_pct)) +
    geom_col(fill = 'green') +
    labs(
        title = "Total amount of cumulative debt between South Korea and Taiwan",
        x = "Year",
        y = "Cumulative Debt"
    )
```

# Question 7-9 Review
# You want to compare the 3-year rolling average of Foreign Direct Investment (fdi_bln) between China and the United States for years starting from 2000.
# Filter & Smooth: Isolate the data for China and the United States for years 2000 and onward.
# Calculate Rolling Averages: Group by country, ensure the data is properly sorted by year, and calculate a new column fdi_3yr_avg using a 3-year rolling window with lag().
# Reshape: Select only year, country, and fdi_3yr_avg, then pivot the data wide so China and United States each have their own column containing these smoothed values.
# Calculate Gap: Create a new column fdi_gap representing China's 3-year rolling average minus the United States' 3-year rolling average (China - United States``).
# Plot: Create a bar chart (geom_col()) showing this FDI rolling average gap (fdi_gap) over time.


```{r}
advanced_macro_data %>%
    filter(country %in% c('United States', 'China')) %>%
    group_by(country) %>%
    arrange(year) %>%
    mutate(rolling_fdi = (fdi_bln +lag(fdi_bln)+ lag(fdi_bln, 2)) / 3) %>%
    filter(year >= 2000) %>%
    select(year, country, rolling_fdi) %>%
    pivot_wider(
        names_from = country,
        values_from = rolling_fdi
    ) %>%
    mutate(fdi_gap = `China` - `United States`) %>%
    ggplot(aes(x = year, y = fdi_gap)) +
    geom_col(fill = 'red')
```





# Question 10
# Take your wide dataset from Question 8 and pivot it back to a long format.
```{r}
advanced_macro_data %>%
    select(year, country, tech_exports_bln) %>%
    pivot_wider(
        names_from = country,
        values_from = tech_exports_bln
    ) %>%
    pivot_longer(
        cols = -year,
        names_to = "country",
        values_to = "tech_exports_bln"
    )
```


# Question 10.1
# Start by pivoting Foreign Direct Investment (fdi_bln) wide across all countries. Then, use pivot_longer() to collapse those country columns back into a long format dataset with columns named country and fdi_bln.


```{r}
advanced_macro_data %>%
    select(year, country, fdi_bln) %>%
    pivot_wider(
        names_from = country,
        values_from = fdi_bln
    ) %>%
    pivot_longer(
        cols = -year,
        names_to = "country",
        values_to = "fdi_bln"
    )
```


# Question 10.2
# Filter the dataset for China, Japan, and South Korea starting from year 2010, and pivot Government Debt (gov_debt_pct) wide. Then, use pivot_longer() to return it back to long format, using cols = c(China, Japan, South Korea) explicitly instead of -year.

```{r}
advanced_macro_data %>%
    filter(country %in% c('China', 'Japan', 'South Korea')) %>%
    select(year, country, gov_debt_pct) %>%
    pivot_wider(
        names_from = country,
        values_from = gov_debt_pct
    ) %>%
    pivot_longer(
        cols = c(`China`, `Japan`, `South Korea`),
        names_to = "country",
        values_to = "gov_debt_pct"
    )
```

# Question 10.3
# Take Unemployment Rate (unemployment_rate) for United States and China, calculate a 3-year rolling average for each, and pivot those rolling averages wide. Finally, use pivot_longer() to collapse the wide rolling average table back into long format with names_to = "country" and values_to = "unemp_3yr_avg".

```{r}
advanced_macro_data %>%
    filter(country %in% c('United States', 'China')) %>%
    group_by(country) %>%
    arrange(year) %>%
    mutate(rolavgun = ((unemployment_rate) + lag(unemployment_rate) + lag(unemployment_rate, 2)) / 3) %>%
    ungroup(country) %>%
    select(year, country, rolavgun) %>%
    pivot_wider(
        names_from = country,
        values_from = rolavgun
    ) %>%
    pivot_longer(
        cols = c(`United States`, China),
        names_to = "country",
        values_to = "rolavgun"
    )
```


# Question 11
# Create a faceted line plot showing gov_debt_pct over time for all countries. However, the y-axis scales must be "free" (independent for each mini-graph) so that countries with lower debt aren't flattened at the bottom of the chart.


```{r}
advanced_macro_data %>%
    ggplot(aes(x = year, y = gov_debt_pct)) +
    geom_line(color = 'blue') +
    facet_wrap(~country, scales = "free_y") +
    labs(
        title = "test",
        x = "Year",
        y = "Government Debt (% of GDP)"
    )
```

# Question 11.1
# Create a faceted line plot showing Foreign Direct Investment (fdi_bln) over time for all countries. Set the y-axis scales to "free_y" so that smaller economies' investment trends are clearly visible alongside larger ones.

```{r}
advanced_macro_data %>%
    ggplot(aes(x = year, y = fdi_bln)) +
    geom_line(color = 'red') +
    facet_wrap(~country, scales = "free_y") +
    labs(
        title = "Second Test",
        x = "Year",
        y = "Foreign Direct Investment (Billions)"
    )
```


# Question 11.2
# Filter the dataset for years 2005 and onward, then create a faceted line plot of Unemployment Rate (unemployment_rate) over time for all countries, using facet_wrap(~ country, scales = "free_y").

```{r}
advanced_macro_data %>%
    filter(year >= 2005) %>%
    ggplot(aes(x = year, y = unemployment_rate)) +
    geom_line(color = 'red') +
    facet_wrap(~country, scales = "free_y") +
    labs(
        title = "Third Test",
        x = "Year",
        y = "Unemployment Rate"
    )
```



# Question 11.3
# Calculate a 3-year rolling average for Technology Exports (tech_exports_bln) for each country (remember to group_by(country) and arrange(year) first!). Then, create a faceted line plot (geom_line()) showing this rolling average over time across all countries using free y-axis scales.


```{r}
advanced_macro_data %>%
    group_by(country) %>%
    arrange(year) %>%
    mutate(rol_avg_teb = (tech_exports_bln + lag(tech_exports_bln, 1) + lag(tech_exports_bln, 2)) / 3) %>%
    ungroup(country) %>%
    ggplot(aes(x = year, y = rol_avg_teb)) +
    geom_line() +
    facet_wrap(~country, scales = "free_y") +
    labs(
        title = "fourth test",
        x = "year",
        y = "rolling average of tech exports"
    )
```

# Question 12
# Plot the fdi_bln of the United States over time. Modify the y-axis so the numbers format automatically with dollar signs (e.g., $100, not 100). Hint: You will need to explore the scales package, which works seamlessly with ggplot.

```{r}
advanced_macro_data %>%
    filter(country == 'United States') %>%
    ggplot(aes(x = year, y = fdi_bln)) +
    geom_line(color = "darkgreen") +
    scale_y_continuous(labels = scales::label_dollar()) +
    labs(
        title = "United States FDI over time",
        x = "Year",
        y = "Foreign Direct Investment (Billions)"
    )
```
# Learned here how to properly add a dollar sign to a chart using the scale_y_continuous function




# Question 12.1
# Plot Technology Exports (tech_exports_bln) for China over time using a line plot (geom_line()). Modify the y-axis so the values automatically format with dollar signs using scales::label_dollar().
```{r}
advanced_macro_data %>%
    filter(country == 'China') %>%
    ggplot(aes(x = year, y = tech_exports_bln))+
    geom_line()+
    scale_y_continuous(labels = scales::label_dollar())
```

# Question 12.2
# Filter the dataset for South Korea and plot its Unemployment Rate (unemployment_rate) over time. Format the y-axis so the numbers display automatically as percentages (e.g., 5% instead of 5) using scale_y_continuous(labels = scales::label_percent(scale = 1)).

```{r}
advanced_macro_data %>%
    filter(country == 'South Korea') %>%
    ggplot(aes(x = year, y = unemployment_rate)) +
    geom_line() +
    scale_y_continuous(labels = scales::label_percent(scale =1))
```

# Question 12.3
# Plot Government Debt (gov_debt_pct) over time for Japan. Modify the y-axis so that numbers are formatted as clean values with comma separators using scale_y_continuous(labels = scales::label_comma()) and add custom labs() titles for clarity.


```{r}
advanced_macro_data %>%
    filter(country == 'Japan') %>%
    ggplot(aes(x = year, y = gov_debt_pct)) +
    geom_line() +
    scale_y_continuous(labels = scales::label_comma())
```



# Question 13
# Create a scatterplot comparing inflation_rate (x) and unemployment_rate (y) for all countries. Remove all rows with NA values before plotting. Change the global theme of the plot to theme_minimal() and move the color legend to the bottom of the chart instead of the right side.

```{r}
advanced_macro_data %>%
    drop_na(inflation_rate, unemployment_rate) %>%
    ggplot(aes(x = inflation_rate, y = unemployment_rate, color = country)) +
    geom_point() +
    theme_minimal() +
    theme(legend.position = "bottom") +
    labs(
        title = "Inflation Rate vs Unemployment Rate",
        x = "Inflation Rate (%)",
        y = "Unemployment Rate (%)",
        color = "Country"
    )
```



# Question 13.1
# Create a scatterplot comparing Government Debt (gov_debt_pct) on the x-axis and Inflation Rate (inflation_rate) on the y-axis for all countries, with points colored by country. Drop rows with NA values in these columns, set the plot theme to theme_light(), and move the legend to the top using theme(legend.position = "top").




```{r}
advanced_macro_data %>%
    drop_na(gov_debt_pct, inflation_rate) %>%
    ggplot(aes(x = gov_debt_pct, y = inflation_rate, color = country)) +
    geom_point() +
    theme_minimal() +
    theme(legend.position = "bottom") +
    scale_y_continuous(labels = scales::label_percent(scale = 1)) +
    labs(
        title = "Comparison of Government Debt to Inflation Rate",
        x = "Government Debt (%)",
        y = "Inflation Rate (%)",
        color = "Country"
    )
```



# Question 13.2
# Filter the data for the United States and China. Create a scatterplot comparing Technology Exports (tech_exports_bln) on the x-axis against FDI (fdi_bln) on the y-axis, colored by country. Use drop_na(), apply theme_bw(), and completely remove the legend using theme(legend.position = "none").





```{r}
advanced_macro_data %>%
filter(country %in% c('United States', 'China')) %>%
drop_na(tech_exports_bln, fdi_bln) %>%
ggplot(aes(x = tech_exports_bln, y = fdi_bln, color = country)) +
geom_point() +
theme_bw() +
theme(legend.position = "bottom") +
labs(
    title = "first test",
    x = "tech exports bln",
    y = "fdi bln"
)
```









# Question 13.3
# Filter the dataset for years 2010 and onward. Drop missing values in fdi_bln and gov_debt_pct, then build a scatterplot comparing fdi_bln (x-axis) and gov_debt_pct (y-axis) colored by country. Apply theme_classic(), move the legend to the bottom, and format the x-axis numbers as currency using scale_x_continuous(labels = scales::label_dollar()).



```{r}
advanced_macro_data %>%
    filter(year >= 2010) %>%
    drop_na(fdi_bln, gov_debt_pct) %>%
    ggplot(aes(x = fdi_bln, y = gov_debt_pct, color = country)) +
    geom_point() +
    theme_classic() +
    theme(legend.position = "bottom") +
    scale_x_continuous(labels = scales::label_dollar()) +
    labs(
        title = "Graph showing fdi to gvt debt",
        x = "fdi (bil)",
        y = "gvt debt (%)"
    )
```



# Question 14 
# Plot unemployment_rate over time for South Korea. Add a vertical dashed line exactly at the year 1997. Add a text annotation next to this line that says "Asian Financial Crisis".




```{r}
advanced_macro_data %>%
    filter(country == 'South Korea') %>%
    ggplot(aes(x = year, y = unemployment_rate)) +
    geom_point() +
    geom_line() +
    geom_vline(xintercept = 1997, linetype = "dashed", color = "red") +
    annotate(
        "text",
        x = 1997.5,
        y = 20,
        label = "Asian Financial Crisis",
        color = "red",
        hjust = 0
    ) +
    labs(
        title = "graph showing asian financial crisis damage.",
        x = "Year",
        y = "Unemployment Rate(%)"
    )
```






# Question 14.1
# Filter the data for the United States and plot Government Debt (gov_debt_pct) over time using geom_line(). Add a vertical dashed red line at the year 2008, and use annotate() to add text at x = 2009 and y = 80 that says "Global Financial Crisis".

```{r}
advanced_macro_data %>%
    filter(country == 'United States') %>%
    ggplot(aes(x = year, y = gov_debt_pct)) +
    geom_line() +
    geom_point() +
    geom_vline(xintercept = 2008, linetype = "solid", color = "red") +
    annotate(
        "text",
        x = 2008.5,
        y = 10,
        label = "Global Financial Crisis",
        color = "red",
        hjust = "red"
    ) +
    labs(
        title = "Graph Showing Effects of the Global Financial Crisis",
        x = "Year",
        y = "Government Debt as a Percent"
    )
```


# Question 14.2
# Plot Inflation Rate (inflation_rate) over time for China. Add a horizontal dashed blue line across the graph at y = 3 using geom_hline(yintercept = 3, linetype = "dashed", color = "blue"). Add a text annotation at x = 2000 and y = 3.5 that says "Inflation Target".


```{r}
advanced_macro_data %>%
    filter(country == 'China') %>%
    ggplot(aes(x = year, y = inflation_rate))+
    geom_line() +
    geom_point() +
    geom_hline(yintercept = 3, linetype = "dashed", color = "blue") +
    annotate(
        "text",
        x = 2000,
        y = 3.5,
        label = "Inflation Target",
        color = "red",
        hjust = 0
    ) +
    labs(
        title = "Graph showing Inflation Target",
        x = "Year",
        y = "Inflation Rate"
    )
```





# Question 14.3
# Filter the dataset for Japan and plot its Unemployment Rate (unemployment_rate) over time using both geom_line() and geom_point(). Add a vertical dashed line at the year 2020 (xintercept = 2020). Annotate the plot at x = 2018 and y = 3.2 with the text "COVID-19 Shock", setting hjust = 1 so the label sits to the left of the dashed line.


```{r}
advanced_macro_data %>%
    filter(country == 'Japan') %>%
    ggplot(aes(x = year, y = unemployment_rate)) +
    geom_line() +
    geom_point() +
    geom_vline(xintercept = 2020, linetype = "dashed", color = "blue") +
    annotate(
        "Text",
        x = 2010,
        y = 7,
        label = "COVID-19 Shock",
        color = "red",
        hjust = 0
    ) +
    labs(
        title = "Graph showing damage of COVID-19",
        x = "Year",
        y = "Unemployment Rate"
    )
```

# Review Question 1
# Before passing to ggplot(), your rolling average pipeline should retain a long, grouped structure with the new calculated column added
# I have an image I am looking at on my end which has the in a horizontal row, countrys at the top left coloumn going down, years are next coloumn, tech exports are the next, and rol avg tech exports are the next so I want that.
# That column only has south korea going from 1995- 1997 then it has the united states under it

```{r}
advanced_macro_data %>%
    filter(year >= 1995, year <= 1997) %>%
    filter(country %in% c('United States', 'South Korea')) %>%
    group_by(country) %>%
    arrange(country, year) %>%
    mutate(rol_avg_teb = (tech_exports_bln + lag(tech_exports_bln) + lag(tech_exports_bln, 2)) / 3) %>%
    ungroup(country) %>%
    select(country, year, tech_exports_bln, rol_avg_teb)
```


# Review Question 2
# When pivoting countries to columns, the dataset collapses years into single rows with countries spread across the horizontal axis:

```{r}
advanced_macro_data %>%
    filter(year >= 2010) %>%
    filter(country %in% c('China', 'Japan', 'South Korea')) %>%
    select(year, country, tech_exports_bln) %>%
    pivot_wider(
        names_from = country,
        values_from = tech_exports_bln
    )

```


# Review Question 3
# When pivoting back from wide, the data expands vertically back into standard key-value pairs:







```{r}
advanced_macro_data %>%
    filter(year >= 2010) %>%
    filter(country %in% c('China', 'Japan', 'South Korea')) %>%
    select(year, country, tech_exports_bln) %>%
    pivot_wider(
        names_from = country,
        values_from = tech_exports_bln
    ) %>%
    pivot_longer(
        cols = -year,
        names_to = "country",
        values_to = "tech_exports_bln"
    )
```






# Review Question 1.1
# Goal: Perform a round-trip reshape (fdi_bln wide by country, then back to long) for China and the United States.
# Final Plot: A line chart comparing FDI over time for both countries with proper currency formatting ($) on the y-axis.


```{r}
advanced_macro_data %>%
    filter(country %in% c('United States', 'China')) %>%
    select(year, country, fdi_bln) %>%
    pivot_wider(
        names_from = country,
        values_from = fdi_bln
    ) %>%
    pivot_longer(
        cols = -year,
        names_to = "country",
        values_to = "fdi_bln"
    ) %>%
    ggplot(aes(x = year, y = fdi_bln, color = country)) +
    geom_line() +
    scale_y_continuous(labels = scales::label_dollar()) +
    labs(
        title = "Comparison of China and United States FDI (billions USD)",
        x = "Year",
        y = "FDI (in Billions)"
    )
```


# Review Question 2.1
# Goal: Calculate a 3-year rolling average of unemployment_rate for all countries (drop NAs).
# Final Plot: A clean small-multiples dashboard (theme_minimal()) showing each country's rolling unemployment trend in its own panel with independent y-axis scales.


```{r}
advanced_macro_data %>%
    arrange(year) %>%
    group_by(country) %>%
    mutate(rlng_avg_unem = (unemployment_rate + lag(unemployment_rate) + lag(unemployment_rate, 2)) / 3) %>%
    ungroup(country) %>%
    ggplot(aes(x = year, y = rlng_avg_unem)) +
    geom_line() +
    geom_point()+
    theme_minimal() +
    facet_wrap(~country, scales = "free_y") +
    scale_y_continuous(labels = scales::label_percent(scale = 1)) +
    labs(
        title = "Three-year rolling averages for unemployment in Major Economies",
        x = "Year",
        y = "Unemployment"
    )
```


# Review Question 3.1
# Goal: Highlight a major macroeconomic event for South Korea.
# Final Plot: A line chart of government debt as a percentage of GDP over time. It must feature a red vertical dashed line at 1997 and a text label reading "IMF Bailout" positioned cleanly at $(1998, 40)$, with the y-axis formatted as percentages.


```{r}
advanced_macro_data %>%
    filter(country == 'South Korea') %>%
    ggplot(aes(x = year, y = gov_debt_pct)) +
    geom_point() +
    geom_line() +
    geom_vline(xintercept = 1997, linetype = "dashed", color = "red") +
    annotate(
        "text",
        x = 1998,
        y = 6,
        label = "IMF Bailout",
        color = "red",
        hjust = 0
    ) +
    labs(
        title = "Graph showing the effects of the IMF Bailout",
        x = "Year",
        y = "Government Debt as a Percentage of GDP"
    )
```



# Review Question 4.1
# Goal: Visualize the Phillips Curve dynamics for data from 2005 onward (drop missing values for both metrics).
# Final Plot: A scatterplot mapping inflation_rate on the x-axis against unemployment_rate on the y-axis, colored by country. Use a light background theme (theme_light()) and position the legend along the bottom.


```{r}
advanced_macro_data %>%
    filter(year >= 2005) %>%
    drop_na(inflation_rate, unemployment_rate) %>%
    ggplot(aes(x = inflation_rate, y = unemployment_rate, color = country)) +
    geom_point() +
    theme_light() +
    theme(legend.position = "bottom")
```

# Review Question 4.2
# Goal: Analyze the relationship between government debt and foreign direct investment for East Asian economies since the 2008 Global Financial Crisis.
# Filtering: Restrict the data to China, Japan, and South Korea for years 2008 and onward. Drop any rows with missing values in gov_debt_pct or fdi_bln.
# Final Plot: A scatterplot mapping gov_debt_pct on the x-axis against fdi_bln on the y-axis, colored by country. Format the y-axis values as currency ($) using scales::label_dollar(), apply theme_minimal(), and move the legend to the top.


```{r}
advanced_macro_data %>%
    filter(year >= 2008) %>%
    filter(country %in% c('China', 'Japan', 'South Korea')) %>%
    drop_na(gov_debt_pct, fdi_bln) %>%
    ggplot(aes(x = gov_debt_pct, y = fdi_bln, color = country)) +
    geom_point() +
    scale_y_continuous(labels = scales::label_dollar()) +
    theme(legend.position = "top") +
    theme_minimal()
```




# Review Question 4.3
# Goal: Evaluate the trade-off between technology exports and economic growth/unemployment dynamics in North America and Asia prior to 2020.
# Filtering: Filter for data before 2020 (year < 2020) for the United States, China, and South Korea. Drop missing values in tech_exports_bln and unemployment_rate.
# Final Plot: A scatterplot mapping tech_exports_bln on the x-axis against unemployment_rate on the y-axis, mapped to color = country. Format the x-axis as dollar values using scale_x_continuous(labels = scales::label_dollar()), apply theme_classic(), and set custom axis/title labels with labs().


```{r}
advanced_macro_data %>%
    filter(year < 2020) %>%
    filter(country %in% c('United States', 'China', 'South Korea')) %>%
    drop_na(tech_exports_bln, unemployment_rate) %>%
    ggplot(aes(x = tech_exports_bln, y = unemployment_rate, color = country)) +
    scale_x_continuous(labels = scales::label_dollar()) +
    scale_y_continuous(labels = scales::label_percent(scale = 1)) +
    #Decided that I would also make sure the unit we are placing in the y category is accurate since we are already doing the same for the x axis
    geom_point() +
    theme_classic() +
    labs(
        title = "Technology Exports to Unemployment Dynamics in North America and Asia",
        x = "Tech Exports (Billions USD)",
        y = "Unemployment Rate (%)"
    )
```

# Review Question 5.1
# Goal: Visualize inflation targets for Japan.Final Plot: A connected point-and-line graph (geom_line + geom_point) of inflation over time. Include a horizontal blue dashed baseline at 2% and a boxed label ("label") reading "BOJ Target" floating at $(2000, 2.5)$.

```{r}
advanced_macro_data %>%
filter(country == 'Japan') %>%
drop_na(year, inflation_rate) %>%
ggplot(aes(x = year, y = inflation_rate)) +
geom_line() +
geom_point() +
geom_hline(yintercept = 2.5, x = 2000, color = "red")
```

# Review Question 6.1
# Goal: Execute a multi-stage pipeline on tech_exports_bln for China, Japan, and South Korea (round-trip pivot $\rightarrow$ 3-year rolling average calculation).Final Plot: A faceted line dashboard showing each country's smoothed tech export trend over time, featuring dollar signs on the y-axes and independent vertical scales.

```{r}
advanced_macro_data %>%
filter(country %in% c('China', 'Japan', 'South Korea')) %>%
select(year, country, tech_exports_bln) %>%
pivot_wider(
    names_from = "country",
    values_from = "tech_exports_bln"
) %>%
pivot_longer(
    cols = -year,
    names_to = "country",
    values_to = "tech_exports_bln"
) %>%
group_by(country) %>%
arrange(year) %>%
mutate(threeyear = (tech_exports_bln + lag(tech_exports_bln) + lag(tech_exports_bln, 2)) / 3) %>%
ungroup() %>%
drop_na(threeyear) %>%
ggplot(aes(x = year, y = threeyear)) +
geom_line() +
facet_wrap(~country, scales = "free_y") +
scale_y_continuous(labels = scales::label_dollar())
```

# Not sure how to do smoothed lines must learn.


# Review Question 7.1
 # Goal: Construct an annotated crisis analysis chart for the United States.Final Plot: A line plot of government debt (% of GDP) using theme_classic(). It must feature:A semi-transparent red shaded rectangle spanning 2007 to 2009.A vertical dashed reference line at 2008.A text annotation reading "Crisis Spike" positioned at $(2009.5, 80)$.

```{r}
advanced_macro_data %>%
    filter(country == 'United States') %>%
    ggplot(aes(x = year, y = gov_debt_pct)) +
    geom_line() +
    theme_classic() +
    geom_vline(xintercept = 2008.5, linetype = "dashed", color = "red") +
    annotate(
        "rect",
        xmin = 2007, xmax = 2009,
        ymin = -Inf, ymax = Inf,
        fill = "red",
        alpha = 0.2
    )
```

# Review Question 8.1
# Goal: Analyze the relationship between government debt and foreign direct investment for East Asian economies since the 2008 Global Financial Crisis.
# Filtering: Restrict the data to China, Japan, and South Korea for years 2008 and onward. Drop any rows with missing values in gov_debt_pct or fdi_bln.
# Final Plot: A scatterplot mapping gov_debt_pct on the x-axis against fdi_bln on the y-axis, colored by country. Format the y-axis values as currency ($) using scales::label_dollar(), apply theme_minimal(), and move the legend to the top.

```{r}
advanced_macro_data %>%
    filter(country %in% c('China', 'Japan', 'South Korea')) %>%
    filter(year >= 2008) %>%
    drop_na(gov_debt_pct, fdi_bln) %>%
    ggplot(aes(x = fdi_bln, y = gov_debt_pct, color = country)) +
    scale_y_continuous(labels = scales::label_dollar()) +
    geom_point() +
    theme_minimal() +
    theme(legend.position = "Top")
```


# Review Question 9.1
# Goal: Evaluate the trade-off between technology exports and economic growth/unemployment dynamics in North America and Asia prior to 2020.
# Filtering: Filter for data before 2020 (year < 2020) for the United States, China, and South Korea. Drop missing values in tech_exports_bln and unemployment_rate.
# Final Plot: A scatterplot mapping tech_exports_bln on the x-axis against unemployment_rate on the y-axis, mapped to color = country. Format the x-axis as dollar values using scale_x_continuous(labels = scales::label_dollar()), apply theme_classic(), and set custom axis/title labels with labs().


```{r}
advanced_macro_data %>%
    filter(country %in% c('United States', 'China', 'South Korea')) %>%
    filter(year < 2020) %>%
    drop_na(tech_exports_bln, unemployment_rate) %>%
    ggplot(aes(x = tech_exports_bln, y = unemployment_rate, color = country)) +
    scale_x_continuous(labels = scales::label_dollar()) +
    theme_classic() +
    geom_point() +
    scale_y_continuous(labels = scales::label_percent()) +
    labs(
        title = "Technology Exports and Unemployment Trade-Off",
        x = "Tech Exports (Billions in USD)",
        y = "Unemployment Rate (%)"
    )
```
    







# Review Question 10.1
# Goal: Practice explicit grid normalization on implicit missing data.
# Task: Filter advanced_macro_data for Taiwan, South Korea, and Japan. Pivot gov_debt_pct wide across countries to force R to create an explicit grid of all years and countries (filling missing year entries with NA). Immediately collapse the grid back into long format using pivot_longer().
# Final Plot: Create a line plot of government debt over time for all three countries, mapping color = country. Use scale_y_continuous(labels = scales::label_percent(scale = 1)) to format the y-axis, and apply theme_minimal().



```{r}
advanced_macro_data %>%
    filter(country %in% c('Taiwan', 'South Korea', 'Japan')) %>%
    select(year, country, gov_debt_pct) %>%
    pivot_wider(
        names_from = "country",
        values_from = "gov_debt_pct"
    ) %>%
    pivot_longer(
        cols = -year,
        names_to = "country",
        values_to = "gov_debt_pct"
    ) %>%
    ggplot(aes(x = year, y = gov_debt_pct, color = country)) +
    scale_y_continuous(labels = scales::label_percent(scale =1)) +
    theme_minimal() +
    geom_point()
```









# Review Question 11.1
# Goal: Perform a round-trip reshape to prepare data for a rolling calculation.
# Task: Take Technology Exports (tech_exports_bln) for China and the United States. Pivot the data wide so each country has its own column, then pivot it back to long format (names_to = "country", values_to = "tech_exports_bln"). Group by country to compute a 3-year rolling average on the restored long dataset.
# Final Plot: Plot this rolling average over time using geom_line(), faceted by country (facet_wrap(~ country, scales = "free_y")) with formatted dollar values (scales::label_dollar()) on the y-axes.



```{r}
advanced_macro_data %>%
    filter(country %in% c('China', 'United States')) %>%
    select(year, country, tech_exports_bln) %>%
    pivot_wider(
        names_from = "country",
        values_from = "tech_exports_bln"
    ) %>%
    pivot_longer(
        cols = -year,
        names_to = "country",
        values_to = "tech_exports_bln"
    ) %>%
    arrange(year) %>%
    group_by(country) %>%
    mutate(threeyear = (tech_exports_bln + lag(tech_exports_bln) + lag(tech_exports_bln, 2)) /3) %>%
    ungroup(country) %>%
    ggplot(aes(x = year, y = threeyear)) +
    geom_line() +
    facet_wrap(~country, scales = "free_y") +
    scale_y_continuous(labels = scales::label_dollar())
```



# Question 15
# Convert the country column into a factor. Relevel this factor so that "South Korea" is explicitly first and "China" is second. Create a boxplot of tech_exports_bln across countries. Because of your factor manipulation, South Korea and China must appear as the first two boxes on the x-axis, regardless of alphabetical order.










Question 2: 

Question 3: 

Question 4: 

Level 2: Advanced Window Functions (Time Series)
In economics, what happened yesterday is just as important as what is happening today. You will need to research window functions in dplyr that look at previous rows.

Question 5: .

Question 6: 

Question 7: 

Level 3: Reshaping Data (Pivoting)
Data rarely comes in the exact shape you need for plotting or regression. You need to master changing data from "long" format to "wide" format, and vice versa.

Question 8: 

Question 9: 

Question 10: 

Level 4: Advanced Visualizations & Scales
Default ggplot settings are rarely presentation-ready for academic papers.

Question 11: 

Question 12: 

Question 13: 

Question 14: 

Question 15: 

Level 5: String Manipulation & Conditional Logic
Real-world data is messy. You must learn to clean text and apply multiple logical conditions simultaneously.

Question 16: Look at the analyst_note column. It is filled with random spaces, varying capitalizations, and underscores. Use the stringr package (part of tidyverse) to clean this column: make everything lowercase, trim all white space from the edges, and replace all underscores with a standard space.

Question 17: Filter the dataset to keep only the rows where the analyst_note contains the word "crisis" or "recession" (ignoring case).

Question 18: Create a new categorical variable called economic_era. If the year is before 1990, label it "Cold War". If the year is between 1990 and 2008, label it "Global Integration". If the year is after 2008, label it "Post-Crisis". Calculate the average gov_debt_pct for the US across these three eras.

Question 19: In a single pipeline, calculate the mean for all numeric columns in the dataset (except year), grouped by country. Hint: Look up how to use the across() function inside summarize().

Question 20 (The Master Pipeline): Write one continuous pipeline that does the following in order:

Drops any rows with NA in inflation_rate.

Filters for years greater than 2000.

Calculates a new column called debt_to_fdi_ratio (gov_debt_pct / fdi_bln).

Groups by country to find the maximum debt_to_fdi_ratio for each nation.

Pipes directly into a ggplot bar chart visualizing this maximum ratio, sorted from highest to lowest.