---
title: "East Asia Economics First Project"
format: html
---
```{r}
library(dplyr)
```



```{r}
macro_data <- read.csv("east_asia_econ_2000_2023.csv")
```

# Question 1
# Load the dataset into an object called macro_data. What command would you use to find the exact number of rows and columns, and what command would you use to print just the names of the variables?


```{r}
dim(macro_data)
```

# Answer: dim

# Question 2. Basic Filtering
# Create a new data frame called china_recent that only includes data for China in years greater than or equal to 2015. How many rows are in this new data frame?

```{r}
china_recent <- macro_data %>%
    filter(country == "China", year >= 2015)
nrow(china_recent)
```

# Question 2 Answer: 9

# Question 3 Logical Operators (OR)
#Suppose you are analyzing economies heavily impacted by early 2000s tech shocks. Filter macro_data to create a dataset containing only observations from Japan OR South Korea. Save it as tech_economies.

```{r}
tech_economies <- macro_data %>%
    filter(country == "Japan" | country == "South Korea")
```

```{r}
View(tech_economies)
```

# Question 4. Logical Operators (AND)
# Filter the dataset for years where a country experienced a recession (defined as gdp_growth < 0) AND had an inflation rate above 2.0%.
```{r}
recession <- macro_data %>%
    filter(gdp_growth < 0, inflation_rate > 2.0)
```

```{r}
View(recession)
```

# Question 4 Answer: No countries experienced a gdp_growth under 0 while having an inflation rate above 2.0%.
# Question 4 meaning: This could be for a variety of reasons, as countries undergo recession the currency becomes more valuable meaning the inflation rate drops since that is inversely related to the demand of money in circulation.

# Question 1.1
# To analyze the impact of the 2008 US financial crisis on East Asian exports, you only need a few columns. Use the select() function to create a dataset named crisis_analysis that contains only the country, year, and gdp_growth columns.

```{r}
crisis_analysis <- macro_data %>%
    select(country, year, gdp_growth)
```

# Question 2.1
# Building on basic filtering, analyze the period immediately following China's entry into the WTO. Filter macro_data for observations where country == "China" and the year is between 2001 and 2006. (You can use logical operators like >= and <= together, or the between() function).

```{r}
China_wto_entry <- macro_data %>%
    filter(country == "China", year <= 2006, year >= 2001)
```

#If I wanted to use the "between" function I could use
```{r}
China_wto_entry <- macro_data %>%
  filter(country == "China", between(year, 2001, 2006))
```

# Question 3.1
# Logical Negation (NOT)Instead of selecting specific economies, suppose you need to exclude Japan from your model. Filter the dataset to include all countries except "Japan" using the != operator. Save it as ex_japan.

```{r}
Not_Japan_data_1 <- macro_data %>%
    filter(country != "Japan")
```

# Question 4.1
# Multiple Logical ConditionsFilter the dataset for years where South Korea or Japan experienced moderate growth ($gdp\_growth \ge 2.0$) while maintaining low inflation ($inflation\_rate \le 1.5$).

```{r}
South_Japan_Growth <- macro_data %>%
    filter(gdp_growth > 2.0, inflation_rate < 1.5, country == "South Korea" | country == "Japan")
```


# Question 1.2 
# Data Dimensions & Counts How complete is your data across economies? Use the count() function on the country column in macro_data to verify exactly how many years of data exist for each economy.

```{r}
macro_data %>%
    count(country)
```

# Question 2.2
# Creating New Variables (Mutate)To gauge aggregate economic distress across your filtered subsets, use mutate() on macro_data to create a new column called misery_index defined as $Misery = Inflation + Unemployment$. (Assume your dataset includes an unemployment_rate column). Save this as macro_expanded.

```{r}
macro_expanded <- macro_data %>%
    mutate(misery_index = inflation_rate + unemployment_rate)
```

# Question 3.2
# Sorting Data (Arrange)To study the peak of regional growth, sort macro_data by gdp_growth in descending order using arrange() and desc(). Which economy and year record the highest growth rate?

```{r}
macro_data %>%
    arrange(desc(gdp_growth))
```

# Question 3.2 Answer: China had the highest growth rate at 14.2% in 2007

# Question 4.2
# Summarizing AggregatesCalculate the overall average economic growth across all observations using summarize() and mean(gdp_growth, na.rm = TRUE).

```{r}
macro_data %>%
    summarize(avg_growth = mean(gdp_growth, na.rm = TRUE))
```
#Learned from this that the summarize command must have mean or sd or min or mix within it, basically it must have a calculation command within it in order to function and that is how it is different from just a standard, View, or mutate command

# Question 1.3
# Grouped SummariesCompare long-term price stability across the region. Group macro_data by country using group_by(), then use summarize() to calculate the average inflation_rate for each individual economy over the 2000–2023 period.

```{r}
macro_data %>%
    group_by(country) %>%
    summarize(mean(inflation_rate))
```